The promise of community-driven preprints in ecology and evolution

### 2024-05-04 ms/ms.qmd
Flipping some % around in the later paragraphs

### *Cautious 'open'-mindedness of research in preprints*
In addition to speeding up dissemination, preprints can also be a useful way to ensure that research remains open and accessible to the research community because the paper is available online no matter what the open access status of the publication ends up being [@bourne2017ten; @vale2015accelerating]. Using the published articles on *EcoEvoRxiv*, we evaluated the open access status of each paper using the R package *roadoi* (vers. `r utils::packageVersion("roadoi")`) to connect to the Unpaywall platform [@jahn2024]. Most of the published articles arising from preprints were open access [`r sum_is_oa$n[2]` (`r sum_is_oa$percent[2]*100`%) [@fig-pubsummary]A]; however, the status of many were unknown [`r sum_is_oa$n[3]` (`r sum_is_oa$percent[3]*100`%)], and `r sum_is_oa$n[1]` (`r sum_is_oa$percent[1]*100`%) were published behind paywalls. The type of open access for a published paper in an open-access journal also varied widely ([@fig-pubsummary]B).

Data and code sharing are also key components of open science [@roche2015public]. In the spirit of 'openness', we expected data and code sharing among preprints to be greater than in many papers published in research journals [@roche2015public; @o2021preferred]. Despite this, we found that `r preprint_data[1,2]` (`r preprint_data[1,"percent"]*100`%) of preprints relying on data on *EcoEvoRxiv* did not share data, and `r preprint_code[1,2]` (`r preprint_code[1,"percent"]*100`%) did not share code. Authors may be reluctant to share data and code for preprints because of the perceived concern that others may acquire and use such data and code prior to publication in a journal. Notably, authors of `r pre_article_data[3,2]` (`r pre_article_data[3,"percent"]*100`%) preprints that did not share data at the preprint stage ultimately did share data when the article was published; authors for `r pre_article_data[2,2]` (`r pre_article_data[2,"percent"]*100`%) preprints never shared data. The same was true for code. Overall, `r pre_article_code[3,2]` (`r pre_article_code[3,"percent"]*100`%) preprints had no open code at the preprint stage but did at the published article stage; authors of `r pre_article_code[2,2]` (`r pre_article_code[2,"percent"]*100`%) preprints did not share code on publication. Code and data-sharing patterns in our sample of preprints are consistent with analyses sharing practices for published articles (e.g., [@o2021preferred]), even for journals with strict public data archiving policies [@roche2015public].

@@

- First, authors are encouraged to share their data and code at the preprint stage. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings [@sarabipour2019value]. Reassuringly, sharing data and code is rarely associated with 'scooping' of research findings [@soeharjono2021reported]. If authors are worried about data being used unintentionally, clear information surrounding its reuse can include in a license (see https://choosealicense.com). Data can also be achived with an embargo on its reuse [@roche2014troubleshooting]. 
- Second, authors need to be aware of the benefits of posting preprints and postprints for constructive feedback [@bourne2017ten]. While it is clear preprints help establish precedence and allow findings to be openly accessible, it is still seems rare that constructive discussions form around preprints in an open forum (e.g. bioRxiv [@anderson2020biorxiv]). Unfortunately, the *EcoEvoRxiv* website does not provide opportunities for discussion given the limitations of the web server at this point in time. As such, we could not accurately assess how much discourse around a given preprint occurs. Clearly, as a community, we need to provide better platforms that document discussions around preprint findings. Such discussions help the authors improve their work and communicate their findings more effectively (when done constructively, of course). One way to facilitate such discussions may be to use open peer reviewing services, such as *Peer Community In* (PCI) or *PubPeer* (see also [@avissar2024recommendations] for more options), to provide feedback on preprints.
- Third, it remains clear that the time between posting a preprint and publication is still quite long. One possible explanation is that preprints are not being sent to suitable journals. Interestingly, out of the `r nrow(data2)` preprints, only `r pci[2,2]` (`r pci[2,3]*100`%) used Peer Community In (PCI) as a preprint peer-review service. Using PCI circumvents editorial decisions without review. As such, it ensures that authors receive faster feedback on a paper and a more realistic recommendation of where such work is likely to be published. Many journals now accept PCI reviews and recommendations when publishing a paper, which may help facilitate faster publication; though more journals could incorperate PCI in thier policies. 
- Finally, while most preprints get seamlessly connected and merged with their published version some remain disconnected. Incorrect cross-linking by indexing platforms (e.g., Google Scholar) can create confusion and lead to frustration among authors. Thankfully, the reasons for unmatched preprints and publications are well understood and easily rectified. They often result from a mismatch between preprint and published metadata (e.g., titles and author details). In fact, it was reasonably common for titles of preprints to change between the preprint and the published article (`r titles[2,2]` or `r titles[2,3]*100`%). We found that mismatched metadata almost always contributed to preprints and published articles not being matched automatically in Google Scholar. We recommend that authors ensure that, when preprints are published in a journal, they update their preprint with the publication DOI, especially if their title has changed. This is quite easy for authors to do on *EcoEvoRxiv*, and it will ensure that the preprint is linked to the published article and that the citations are merged. It will also create a link between the final preprint and the published paper so that, no matter the open access status of the publication, the research findings will remain openly accessible, saving costs for authors and improving the visibility and use of research.

>>>>

### *Cautious 'open'-mindedness of research in preprints*
In addition to speeding up dissemination, preprints can also be a useful way to ensure that research remains open and accessible to the research community because the paper is available online no matter what the open access status of the publication ends up being [@bourne2017ten; @vale2015accelerating]. Using the published articles on *EcoEvoRxiv*, we evaluated the open access status of each paper using the R package *roadoi* (vers. `r utils::packageVersion("roadoi")`) to connect to the Unpaywall platform [@jahn2024]. Most of the published articles arising from preprints were open access [`r sum_is_oa$percent[2]*100`% (*n* = `r sum_is_oa$n[2]`); [@fig-pubsummary]A]; however, the status of many were unknown [`r sum_is_oa$percent[3]*100`% (*n* = `r sum_is_oa$n[3]`)], and `r sum_is_oa$percent[1]*100`% (*n* = `r sum_is_oa$n[1]`) were published behind paywalls. The type of open access for a published paper in an open-access journal also varied widely ([@fig-pubsummary]B).

Data and code sharing are also key components of open science [@roche2015public]. In the spirit of 'openness', we expected data and code sharing among preprints to be greater than in many papers published in research journals [@roche2015public; @o2021preferred]. Despite this, we found that `r preprint_data[1,"percent"]*100`%  (*n* = `r preprint_data[1,2]`) of preprints relying on data on *EcoEvoRxiv* did not share data, and `r preprint_code[1,"percent"]*100`% (*n* = `r preprint_code[1,2]`) did not share code. Authors may be reluctant to share data and code for preprints because of the perceived concern that others may acquire and use such data and code prior to publication in a journal. Notably, authors of `r pre_article_data[3,"percent"]*100`% (*n* = `r pre_article_data[3,2]`) preprints that did not share data at the preprint stage ultimately did share data when the article was published; authors for `r pre_article_data[2,"percent"]*100`% (*n* = `r pre_article_data[2,2]`) preprints never shared data. The same was true for code. Overall, `r pre_article_code[3,"percent"]*100`% (*n* = `r pre_article_code[3,2]`) preprints had no open code at the preprint stage but did at the published article stage; authors of `r pre_article_code[2,"percent"]*100`% (*n* = `r pre_article_code[2,2]`) preprints did not share code on publication. Code and data-sharing patterns in our sample of preprints are consistent with analyses sharing practices for published articles (e.g., [@o2021preferred]), even for journals with strict public data archiving policies [@roche2015public].

@@

- First, authors are encouraged to share their data and code at the preprint stage. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings [@sarabipour2019value]. Reassuringly, sharing data and code is rarely associated with 'scooping' of research findings [@soeharjono2021reported]. If authors are worried about data being used unintentionally, clear information surrounding its reuse can include in a license (see https://choosealicense.com). Data can also be achived with an embargo on its reuse [@roche2014troubleshooting]. 
- Second, authors need to be aware of the benefits of posting preprints and postprints for constructive feedback [@bourne2017ten]. While it is clear preprints help establish precedence and allow findings to be openly accessible, it is still seems rare that constructive discussions form around preprints in an open forum (e.g. bioRxiv [@anderson2020biorxiv]). Unfortunately, the *EcoEvoRxiv* website does not provide opportunities for discussion given the limitations of the web server at this point in time. As such, we could not accurately assess how much discourse around a given preprint occurs. Clearly, as a community, we need to provide better platforms that document discussions around preprint findings. Such discussions help the authors improve their work and communicate their findings more effectively (when done constructively, of course). One way to facilitate such discussions may be to use open peer reviewing services, such as *Peer Community In* (PCI) or *PubPeer* (see also [@avissar2024recommendations] for more options), to provide feedback on preprints.
- Third, it remains clear that the time between posting a preprint and publication is still quite long. One possible explanation is that preprints are not being sent to suitable journals. Interestingly, out of the `r nrow(data2)` preprints, only `r pci[2,3]*100`% (*n* = `r pci[2,2]`) used Peer Community In (PCI) as a preprint peer-review service. Using PCI circumvents editorial decisions without review. As such, it ensures that authors receive faster feedback on a paper and a more realistic recommendation of where such work is likely to be published. Many journals now accept PCI reviews and recommendations when publishing a paper, which may help facilitate faster publication; though more journals could incorperate PCI in thier policies. 
- Finally, while most preprints get seamlessly connected and merged with their published version some remain disconnected. Incorrect cross-linking by indexing platforms (e.g., Google Scholar) can create confusion and lead to frustration among authors. Thankfully, the reasons for unmatched preprints and publications are well understood and easily rectified. They often result from a mismatch between preprint and published metadata (e.g., titles and author details). In fact, it was reasonably common for titles of preprints to change between the preprint and the published article [`r titles[2,3]*100`% (*n* = `r titles[2,2]`)]. We found that mismatched metadata almost always contributed to preprints and published articles not being matched automatically in Google Scholar. We recommend that authors ensure that, when preprints are published in a journal, they update their preprint with the publication DOI, especially if their title has changed. This is quite easy for authors to do on *EcoEvoRxiv*, and it will ensure that the preprint is linked to the published article and that the citations are merged. It will also create a link between the final preprint and the published paper so that, no matter the open access status of the publication, the research findings will remain openly accessible, saving costs for authors and improving the visibility and use of research.
